---
title: "MotrpacBicQC: Proteomics QC"
date: "`r Sys.Date()`"
output:
  rmdformats::downcute:
    code_folding: show
    self_contained: true
    thumbnails: false
    lightbox: true
pkgdown:
  as_is: true
  
vignette: >
  %\VignetteEngine{knitr::knitr}
  %\VignetteIndexEntry{MotrpacBicQC: Proteomics QC}
  %\usepackage[UTF-8]{inputenc}
---


## Installation

First, download and install R and RStudio:

- [R](https://mirror.las.iastate.edu/CRAN/) 
- [RStudio](https://rstudio.com/products/rstudio/download/) (free version)

Then, open RStudio and install the `devtools` package

```{r eval = FALSE}
install.packages("devtools")
```

Finally, install the `MotrpacBicQC` package

```{r eval = FALSE}
library(devtools)
devtools::install_github("MoTrPAC/MotrpacBicQC", build_vignettes = TRUE)
```


## Usage

Load the library

```{r setup}
library(MotrpacBicQC)
```

And run any of the following tests to check that the package 
is correctly installed and it works. For example:

```{r eval = FALSE}
# Just copy and paste in the RStudio terminal
test <- check_ratio_proteomics(df_ratio = metadata_metabolites_named, 
                               isPTM =  TRUE)
test <- check_rii_proteomics(df_rri = metadata_metabolites_named, 
                             isPTM =  TRUE)
test <- check_vial_metadata_proteomics(df_vm = metadata_metabolites_named)

```

which should generate the following outputs:

```
- (-) The following required columns are missed: ptm_id, protein_id, gene_symbol, entrez_id
- (-) The following required columns are missed: protein_id, sequence, ptm_id, ptm_peptide, gene_symbol, entrez_id, confident_score, confident_site
- (-) The following required columns are missed: vial_label, tmt_plex, tmt11_channel
```

## How to test your datasets

Two approaches available:

### Check full `RESULTS_YYYYMMDD` folder (recommended)

```
|-- PASS1B-06
|   |-- T55
|   |   |-- PROT_PH
|   |   |   `-- BATCH1_20200312
|   |   |       |-- RAW_20200312
|   |   |       |   |-- 01MOTRPAC_PASS1B-06_T55_PH_PN_20191231
|   |   |       |   |   |-- 01MOTRPAC_PASS1B-06_T55_PH_PN_201912319_MANIFEST.txt
|   |   |       |   |   `-- 01MOTRPAC_PASS1B-06_T55_PH_PN_201912319_TMTdetails.txt
|   |   |       |   |-- 02MOTRPAC_PASS1B-06_T55_PH_PN_20191231
|   |   |       |   |   |-- 02MOTRPAC_PASS1B-06_T55_PH_PN_201912319_MANIFEST.txt
|   |   |       |   |   `-- 02MOTRPAC_PASS1B-06_T55_PH_PN_201912319_TMTdetails.txt
|   |   |       |   |-- 03MOTRPAC_PASS1B-06_T55_PH_PN_20191231
|   |   |       |   |   |-- 03MOTRPAC_PASS1B-06_T55_PH_PN_201912319_MANIFEST.txt
|   |   |       |   |   `-- 03MOTRPAC_PASS1B-06_T55_PH_PN_201912319_TMTdetails.txt
|   |   |       |   |-- 04MOTRPAC_PASS1B-06_T55_PH_PN_20191231
|   |   |       |   |   |-- 04MOTRPAC_PASS1B-06_T55_PH_PN_201912319_MANIFEST.txt
|   |   |       |   |   `-- 04MOTRPAC_PASS1B-06_T55_PH_PN_201912319_TMTdetails.txt
|   |   |       |   |-- 05MOTRPAC_PASS1B-06_T55_PH_PN_20191231
|   |   |       |   |   |-- 05MOTRPAC_PASS1B-06_T55_PH_PN_201912319_MANIFEST.txt
|   |   |       |   |   `-- 05MOTRPAC_PASS1B-06_T55_PH_PN_201912319_TMTdetails.txt
|   |   |       |   `-- 06MOTRPAC_PASS1B-06_T55_PH_PN_20191231
|   |   |       |       |-- 06MOTRPAC_PASS1B-06_T55_PH_PN_201912319_MANIFEST.txt
|   |   |       |       `-- 06MOTRPAC_PASS1B-06_T55_PH_PN_201912319_TMTdetails.txt
|   |   |       |-- RESULTS_20200909
|   |   |       |   |-- MOTRPAC_PASS1B-06_T55_PH_PN_20200909_results_RII-peptide.txt
|   |   |       |   |-- MOTRPAC_PASS1B-06_T55_PH_PN_20200909_results_ratio.txt
|   |   |       |   `-- MOTRPAC_PASS1B-06_T55_PH_PN_20200909_vial_metadata.txt
|   |   |       `-- file_manifest_20200910.csv
```


Run test on the full submission. For that, run the following command:

```{r eval = FALSE}
validate_proteomics(input_results_folder = "/full/path/to/RESULTS_YYYYMMDD", 
                    cas = "your_site_code")

# in the example above...
validate_proteomics(input_results_folder = "/full/path/to/PASS1B-06/T55/PROT_PH/BATCH1_20200312/RESULTS_20200909", 
                    cas = "pnnl",
                    isPTM = TRUE,
                    return_n_issues = FALSE)

```

**cas** is one of the followings:

- "broad_prot" = Broad Proteomics
- "pnnl"       = PNNL


### Check individual files

- Check ratio resuls file

```{r eval = FALSE}
# Open the ratio results file

proteomics_ratio_results <- read.delim(file = "/path/to/your/file", stringsAsFactors = FALSE)

check_ratio_proteomics(df_ratio = proteomics_ratio_results, 
                       isPTM = TRUE, 
                       printPDF = FALSE)

```

- Check proteomics reporter ion intensity file:

```{r eval = FALSE}
# Open your files
proteomics_ratio_results <- read.delim(file = "/path/to/your/file", stringsAsFactors = FALSE)

check_rii_proteomics(df_rri = proteomics_ratio_results, cas = "your_side_id")
```

- Check proteomics vial metadata file

```{r eval = FALSE}
# Open your files
vial_metadata <- read.delim(file = "/path/to/your/file", stringsAsFactors = FALSE)

check_vial_metadata_proteomics(df_vm = vial_metadata)
```



## Help

Additional details for each function can be found by typing, for example:

```{r eval=FALSE}
?check_vial_metadata_proteomics
```

Need extra help? Please, [submit an issue here](https://github.com/MoTrPAC/MotrpacBicQC/issues) 
providing as many details as possible.

